---
title: "Gov 50 Final Project: GDP Growth and Democratization in Kyrgyzstan"
author: "Roxana Joyce Chen"
description: "Does GDP growth facilitate democratization? In this study, I plan to examine the extent to which GDP growth begets democratic progress. I hypothesize that GDP growth allows democratic progress to occur through factors such as increased governmental capacity, heightened foreign investment, strengthened economic and political institutions, and a more stable middle class."
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

Does GDP growth facilitate democratization?  In this study, I plan to examine the extent to which GDP growth begets democratic progress. I hypothesize that GDP growth allows democratic progress to occur through factors such as increased governmental capacity, heightened foreign investment, strengthened economic and political institutions, and a more stable middle class. My explanatory variable of interest is the Kyrgyz GDP growth rate from its founding in 1991 to present, according to [World Bank data](https://data.worldbank.org/indicator/NY.GDP.MKTP.KD.ZG?locations=KG). My outcome variable is the state of liberal democracy in Kyrgyzstan in each associated year, given by VDem data capturing [Kyrgyzstan’s annual Liberal Democracy Index (LDI)](https://www.v-dem.net/data_analysis/CountryGraph/). The LDI is measured one a scale from 0 to 1, and is an aggregate rating of a country’s electoral democracy levels, civil liberty protections, degrees of legal and judicial independence, and governance accountability mechanisms. If I observe greater scores on the LDI in periods of Kyrgyz economic growth, this would support my hypothesis that GDP growth is positively associated with democratic hprogress; on the other hand, if LDI scores were to drop during periods of economic growth, this evidence would contradict my hypothesis. When I ultimately run my regression, I would need a positive, statistically significant coefficient to corroborate my hypothesis.


```{r}
library(tidyverse)
library(broom)
library(ggplot2)
library(modelr)

#Loading and joining data
ldi <- read_csv("Kyrgyzstan.csv")
gdp_growth <- read_csv("wbdatacleaned.csv")
kyrgyz <- ldi |>
 left_join(gdp_growth)

#Visualization of GDP growth over time
gdp_growth_graph <- 
  ggplot(data = kyrgyz, 
          mapping = aes(x = `Year`, 
                        y = `GDP growth (annual %)`)) +
  geom_line() +
  labs(x = "Year", 
       y = "GDP growth (Annual %)", 
       title = "Annual GDP Growth in Kyrgyzstan, 1991-2021")
gdp_growth_graph

#Visualization of LDI over time
ldi_graph <- 
  ggplot(data = kyrgyz, 
          mapping = aes(x = `Year`, 
                        y = `Liberal Democracy Index`)) +
  geom_line() +
  labs(x = "Year", 
       y = "Liberal Democracy Index Score", 
       title = "Liberal Democracy Index Score in Kyrgyzstan, 1991-2021")
ldi_graph

#Linear regression 
regression <- lm(`Liberal Democracy Index` ~ `GDP growth (annual %)`, 
                 data = kyrgyz)
regression
summary(regression)

#Coefficients
coef(regression)

#Graphing regression
ggplot(regression, aes(x = `GDP growth (annual %)`, 
                       y = `Liberal Democracy Index`)) + 
  geom_point(color = "steelblue1") + 
  labs(x = "GDP growth (Annual %)", 
       y = "Liberal Democracy Index", 
       title = "Liberal Democracy Index vs. Annual GDP Growth in Kyrgyzstan,
       1991-2021" ) +
#geom vline, hline
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "indianred1", 
              size = 1.5)
```

 (1) an introduction where you introduce the research question and hypothesis and briefly describe why it is interesting

```{r}

```

 
 (2) a data section that briefly describes the data source, describes how the key dependent and independent variables are measured (e.g., a survey, statistical model, or expert coding), and also produces a plot that summarizes the dependent variable; note if your research design is cross-sectional (most projects will be of this type) or one of the other designs we discussed (randomized experiment, before-and-after, differences-in-differences).
 
```{r}
#not causal; association. observing relationship
```
 
 (3) a results section that contains a scatterplot, barplot, or boxplot of the main relationship of interest and output for the main regression of interest; interpret (in plain English) the main coefficient of interest in your regression. You should also comment on the statistical significance of the estimated coefficient and whether or not you believe the coefficient to represent a causal effect. 
 
```{r}

#LDI and GDP alone
#if LDI 0, gdp would be.
#if GDP growth 1 pp increase, LDI 
#On average, LDI increases by x if GDP growth rate increases by 1 percentage point [contingent on holding [[other variable]]] fixed

#statistically significant.

#Regression with important confounders: LDI, GDP, Aid, Revolutions, Inequality (and explain why important). 
#interpretations of coefficient
#does association go away?
#is it still significant?
  
#run a separate regression with revolution

```
 
 (4) a brief (one paragraph) concluding section that summarizes your results, assesses the extent to which you find support for your hypothesis, describes limitations of your analysis and threats to inference, and states how your analysis could be improved (e.g., improved data that would be useful to collect).
 
```{r}
#Results Summary
#Hypothesis Support
#Limitations and Threats
#Improvements

#is positive assoc, but can it be explained by other things - if put other variables in, does association go away?
```
 
 